---
title: "analiza"
author: "Mikołaj Data"
date: "2024-05-20"
output: html_document
---

```{r}
install.packages("knitr")
install.packages("treemap")
install.packages("readr")
install.packages("tidyverse")
install.packages("readr")
install.packages('cluster')
install.packages('factoextra')
install.packages('randomForest')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read the data
Data source: https://www.kaggle.com/datasets/teejmahal20/airline-passenger-satisfaction

```{r}
library(readr)
dat <- read_csv("train.csv")
dat <- dat[,-2:0]
dat <- na.omit(dat)
dat
```
# Prepare data and visualize
In case of you wondering this is source of charts ideas:
https://r-graph-gallery.com/index.html


```{r}
library("RColorBrewer")
library("ggplot2")

dat$Gender <- as.factor(dat$Gender)
coul <- brewer.pal(6, "Set2") 
barplot(summary(dat$Gender), col=coul)
```

```{r}
dat$`Customer Type` <-as.factor(dat$`Customer Type`)
barplot(summary(dat$`Customer Type`), col=coul)
```

```{r}
dat$satisfaction <- as.factor(dat$satisfaction)
barplot(summary(dat$satisfaction), col=coul)
```

```{r}
library(treemap)
dat$Class <- as.factor(dat$Class)

values <- c(sum(dat$Class == "Eco"), 
            sum(dat$Class == "Eco Plus"), 
            sum(dat$Class == "Business"))

groups <- c("Eco", "Eco Plus", "Business")
df <- data.frame(groups, values)
treemap(df, index="groups",
        vSize="values", type="index")
```

```{r}
library(dplyr)
colnames(dat)
names<-c("Inflight wifi service", "Ease of Online booking", "Food and drink", "Seat comfort",
         "On-board service", "Baggage handling", "Departure Delay in Minutes",
         "Departure/Arrival time convenient", "Gate location", "Online boarding",
         "Inflight entertainment", "Leg room service", "Checkin service",
         "Arrival Delay in Minutes")

lapply(dat[names] %>% na.omit(), mean)
```

```{r}
dat$`Type of Travel` <- as.factor(dat$`Type of Travel`)
barplot(summary(dat$`Type of Travel`), col=coul)
```

# Plot some correlations
```{r}
plot(dat[1:200,(ncol(dat)-5):ncol(dat)-1], cex=0.5)
```
```{r}
heatmap(as.matrix(dat[1:100,(ncol(dat)-16):ncol(dat)-1]),Colv = NA, Rowv = NA, scale="column")
```
# Let's try to divide a satisfaction label on 3 items "dissatisfied", "neutral" and "satisfied"
```{r}
df <- dat[, -(1:6)]
df <-df[,1:16]
df <- scale(df)
km.out <- kmeans(df, centers=3)
clustered <- cbind(dat, cluster=km.out$cluster)
```

Evaluate results
```{r}
library(purrr)
satisfied <-  sum(clustered$satisfaction == "satisfied")
dissatisfied_or_neutral <- sum(clustered$satisfaction == "neutral or dissatisfied")
satisfied_clustered_properly <- sum(clustered$satisfaction == "satisfied" 
                                    & clustered$cluster == 1)
dissatisfied_clustered_properly <- sum(clustered$satisfaction == "neutral or dissatisfied" 
                                       & clustered$cluster == 3)
neutral_clustered_properly <- sum(clustered$satisfaction == "neutral or dissatisfied" & clustered$cluster == 2)

print(paste("Satisfied clustered properly: ", satisfied_clustered_properly/satisfied))
print(paste("Dissatisfied neutral", (dissatisfied_clustered_properly
+neutral_clustered_properly)/dissatisfied_or_neutral))
```
Results: Disapointed but not surprised :( clustering in this way is a little bit too random.

```{r}
clustered$cluster[clustered$satisfaction == "satisfied" & clustered$cluster > 1] <- 1
clustered$cluster[clustered$satisfaction == "neutral or dissatisfied" & clustered$cluster == 1] <- 4
clustered <- clustered[, -which(names(clustered) == "satisfaction")]
```

```{r}
clustered$cluster <- as.factor(clustered$cluster)
barplot(summary(clustered$cluster), col=coul)
```
# Diffrent aproach to the problem
https://letsdatascience.com/binning/
```{r}
df <- dat[, -(1:6)]
df <-df[,1:14]
df_with_means <- cbind(dat, rank_mean=apply(df, 1, mean))
df_with_means
```
In this point we can see that passanger with higher mean are most likely to be satisfied :)
Calculate this ;)
```{r}
print(mean(df_with_means[df_with_means$satisfaction=="satisfied",]$rank_mean))
print(mean(df_with_means[df_with_means$satisfaction=="neutral or dissatisfied",]$rank_mean))
```

```{r}
sample <- sample(c(TRUE, FALSE), nrow(df), replace=TRUE, prob=c(0.7,0.3))
train  <- df[sample, ]
test   <- df[!sample, ]
```